Soss

### 2019-12-30 paper/paper.tex
typo
works of GG.jl by Taine Zhao

@@

Soss began with a goal of representing models with continuous, fixed-dimensionality parameter spaces, inspied by \emph{Stan} \cite{stan:2017}.

>>>>

\todo[inline]{Restructure my writing?}

Besides the purely static code generation via regular macros happening in parsing time, Soss heavily uses a mechanism of "zero-cost" runtime code generation originated by Julia's generated functions \cite{bezanson2012julia}, a.k.a staged functions in the original paper and earlier versions of Julia Programming Language.

The generated functions provide the capability of performing code generation during type inference, generating programs computed by the body of function(a.k.a, generator), once and only once for each combination of argument types.

Further, Julia enables type inference and compiler optimizations equivalent to non-runtime ones in runtime for the callsites of generated functions, hence we gain runtime code generation without losing performance.

To take advantage of this, Soss designs a system to encode sufficient information for generating actual codes, into the types, or objects that can be dispatched like types, which allows the use of generated functions here.

Unfortunately, in current stage, there're some implementation restrictions to Julia's generated functions, and generated functions cannot generate arbitrary Julia codes,
where some advanced programming constructions are missing, such as generators and function-related stuffs like closures(including functions with no free variables), multiple dispatch, etc.

To address this, we introduced the works of easing the restrictions of Julia generated functions,
\todo[inline]{cite GG here?}
which allow us to generate closures from generated functions, and make the programs sufficiently powerful.

Programs that can be generated by Soss are turing complete, as all constructs of Lambda Calculus can be generated.

\begin{lstlisting}[language = Julia]
  abstract type LCTerm end
  struct Lam{Arg, Body <: LCTerm} <: LCTerm end
  struct App{Fn<:LCTerm, Arg<:LCTerm} <: LCTerm end
  struct Var{N} <: LCTerm end

  cg(::Type{Lam{Arg, Term}}) where {Arg, Term} = 
    :($Arg -> $(cg(Term)))
  cg(::Type{App{Fn, Arg}}) where {Fn, Arg} = 
    :($(cg(Fn))($(cg(Arg))))
  cg(::Type{Var{N}}) where N = N

  function tolc(expr)
    @match expr begin
      :($f($arg)) => App{tolc(f), tolc(arg)}
      :($x => $y) => Lam{x, tolc(y)}
      a::Symbol => Var{a}
    end
  end

  @gg function lc(::Type{Term}) where Term <: LCTerm
      cg(Term)
  end

  macro lc(term)
    lc(tolc(term))
  end

  x = Core.Box()
  y = Core.Box()
  z = Core.Box()
  f(x) = (x, z)
  @assert (@lc x => x)(x) === (@lc y => y)(x)
  @assert (@lc f => x => f(x))(f)(y) === (y, z)
  \end{lstlisting}

@@

Soss began with a goal of representing models with continuous, fixed-dimensionality parameter spaces, inspired by \emph{Stan} \cite{stan:2017}.
